---
title: "Clustering Template"
author: "Sophia Gao"
date: "February 16, 2015"
output: html_document
---

```{r setUp,render=FALSE}
# libraries
library(rattle) # delete

library(dplyr)
library(corrplot)
library(NbClust)
library(Hmisc)

# project path
folderPath <- "~/Projects/predictiveModelingTempatesR/clustering"
setwd(folderPath)
dataRead <- "test.csv"
dataWrite <- "output.csv"
```

```{r import Data}
# rawData = read.csv(dataRead, header=TRUE, sep = ",")
data(wine, package="rattle")
rawData <- wine[-1]

# dplyr
rawData <- tbl_df(rawData)

# variable name list
varNames <- names(rawData)
varNames
keysVarNames <- NULL
predVarNames = varNames[!(varNames %in% c(keysVarNames))]
charVarNames <- predVarNames[sapply(rawData[,predVarNames],is.character)]
numVarNames <- setdiff(predVarNames,charVarNames)
```

## Data Quality Investigation
```{r dataQuality,echo=TRUE}
# Duplicated "ID"s(Rows)
# Group by "ID"s
# Columns you want to group by
grp_cols <- predVarNames
# Convert character vector to list of symbols
dots <- lapply(grp_cols,as.symbol)
# Perform frequency counts
counts <- rawData %>% group_by_(.dots=dots) %>% summarise(count = n())
describe(counts$count)

# Summary of Variables
describe(rawData)

# missing rate of all the variables
missingRate <- function(x){
    sum(is.na(x))/length(x)
}

missingRateV = sapply(rawData,missingRate)
missingRateV[missingRateV>0.01]
```

```{r pre-processing, include=FALSE}
cleanData <- scale(rawData)
```


K-Means Clustering
```{r}
wssplot <- function(data, nc=15, seed=1234){
               wss <- (nrow(data)-1)*sum(apply(data,2,var))
               for (i in 2:nc){
                    set.seed(seed)
                    wss[i] <- sum(kmeans(data, centers=i)$withinss)}
                plot(1:nc, wss, type="b", xlab="Number of Clusters",
                     ylab="Within groups sum of squares")
               }

```

```{r}
wssplot(cleanData)
```
There is a distinct drop in within groups sum of squares when moving from 1 to 3 clusters. After three clusters, this decrease drops off, suggesting that a 3-cluster solution may be a good fit to the data. 

NbClust
NbClust package provides 30 indices for determining the number of clusters and proposes to user the best clustering scheme from the different results obtained by varying all combinations of number of clusters, distance measures, and clustering methods.

```{r}
set.seed(1234)
nc <- NbClust(cleanData, min.nc = 2, max.nc = 15, method = "kmeans")
table(nc$Best.n[1,])
barplot(table(nc$Best.n[1,]), 
          xlab="Numer of Clusters", ylab="Number of Criteria",
          main="Number of Clusters Chosen by 26 Criteria")
```
14 of 24 criteria provided by the NbClust package suggest a 3-cluster solution. 

A final cluster solution
```{r}
set.seed(1234)
fit.km <- kmeans(cleanData, 3, nstart=25)                           
fit.km$size
fit.km$centers  

aggregate(cleanData, by=list(cluster=fit.km$cluster), mean)
```
